Transition path sampling and the calculation of free energies of enzymatic reactions

### 2022-03-25 tex/manuscript.tex
added some more references

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcb,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
%\usepackage{graphicx}
\usepackage{graphicx}
\usepackage{caption}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Atomistic level study of the mechanism of enzyme catalysis is necessary
to understand the efficiency of enzymes as well as to device strategies for inhibiting 
their activity. 
Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008} 
\section{Theory}
\subsection{Transition path sampling}
%Transition path sampling is an importance sampling in the trajectory space.
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is run for a total 
time of $\mathcal{T}$ the number of time slices is given by $L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
\begin{equation}
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
\end{equation}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\begin{equation}
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
\end{equation}
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\begin{equation}
\rho(z_0) = \exp(-\beta H(z_0))/Z
\end{equation}
where 
\begin{equation}
Z = \int dz \exp(-\beta H(z)) 
\end{equation}
is the canonical partition function.

@@

\section{Results and discussion}
%\begin{figure}
%\centering
%\includegraphics[scale=0.2 ]{figures/ts253.png}
%\caption{Enzymatic reaction in human Mat2a enzyme}
%  \label{fgr:mat2a_reaction}
%\end{figure}
\section{Adenosine deaminase reaction coordinates and free energy}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/reactant-state.png}
\caption{Reactant state}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/int1.png}
\caption{Intermediate state}
\label{fig:minipage2}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/product-state.png}
\caption{Product state}
\label{fig:minipage1}
\end{minipage}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcb,manuscript=article,reprint]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Atomistic level study of the mechanism of enzyme catalysis is necessary
to understand the efficiency of enzymes as well as to device strategies for inhibiting 
their activity. 
Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
is an unbiased monte carlo method in the trajectory space which is used for the calculation 
of reaction coordinates and transition states of complex chemical reactions. 
TPS uses reactive trajectories which connect a reactant state to a product state and these 
trajectories are rare in molecular dynamics simulations because of the energy 
barriers separating the reactants and the products. To overcome these energy barriers and 
to obtain ergodic sampling several methods have been developed over the years such as 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, etc. These methods rely on applying bias
potentials for the system to overcome energy barriers and hence require the knowledge of 
the reaction coordinates.

TPS based free energy simulations have been attempted previously by Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}

Human Mat2a enzyme and plasmodium vivax adenosine deaminase \cite{Luo07JAmChemSoc129p8008} 
\section{Theory}
\subsection{Transition path sampling}
%Transition path sampling is an importance sampling in the trajectory space.
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is run for a total 
time of $\mathcal{T}$ the number of time slices is given by $L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
\begin{equation}
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
\end{equation}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probablility can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\begin{equation}
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
\end{equation}
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\begin{equation}
\rho(z_0) = \exp(-\beta H(z_0))/Z
\end{equation}
where 
\begin{equation}
Z = \int dz \exp(-\beta H(z)) 
\end{equation}
is the canonical partition function.

@@

\section{Results and discussion}
%\begin{figure}
%\centering
%\includegraphics[scale=0.2 ]{figures/ts253.png}
%\caption{Enzymatic reaction in human Mat2a enzyme}
%  \label{fgr:mat2a_reaction}
%\end{figure}
\section{Adenosine deaminase reaction coordinates and free energy}
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.25]{./figures/ada-protein.png}
\end{figure}

\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/reactant-state.png}
\caption{Reactant state}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/int1.png}
\caption{Intermediate state}
\label{fig:minipage2}
\end{minipage}
\quad
\begin{minipage}[b]{0.32\linewidth}
\includegraphics[scale=0.3]{figures/product-state.png}
\caption{Product state}
\label{fig:minipage1}
\end{minipage}
\end{figure}
